---
title: "Markdown_of_UN_Data_pt2"
author: "Isaac"
date: "2024-10-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(tidyverse)
```

```{r}
gapminder_data <- read_csv("gapminder_data.csv")

summary(gapminder_data)
```

```{r}
gapminder_data <- read_csv("gapminder_data.csv")

summary(gapminder_data)

gapminder_data %>%
  summarize(averageLifeExp = mean(lifeExp))

gapminder_data_summarized <- gapminder_data %>%
  summarize(averageLifeExp=mean(lifeExp))

gapminder_data_summarized

gapminder_data
```


```{r}

gapminder_data %>%  
  summarize(recent_year = max(year))

gapminder_2007 <- gapminder_data %>%
  filter(year == 2007) %>%
  summarize(average=mean(lifeExp))

gapminder_data %>%
  summarize(first_year=min(year))

gapminder_data %>%
  filter(year == 1952) %>%
  summarize(average_gdp=mean(gdpPercap))

gapminder_data %>%
  group_by(year) %>%
  summarize(average=mean(lifeExp))

gapminder_data %>%
  group_by(continent) %>%
  summarize(average=mean(lifeExp))
```



# all the continents min, max and average life exp. for all of the years and 1952(oldest) 2007 (most recent)

```{r}
life_exp_by_continent <- gapminder_data %>%
  group_by(continent) %>%
  summarize(min=min(lifeExp), average=mean(lifeExp), max=max(lifeExp))

life_exp_by_continent_1952 <- gapminder_data %>%
  group_by(continent) %>%
  filter(year == 1952) %>%
  summarize(min=min(lifeExp), average=mean(lifeExp), max=max(lifeExp))

life_exp_by_continent_2007 <- gapminder_data %>%
  group_by(continent) %>%
  filter(year == 2007) %>%
  summarize(min=min(lifeExp), average=mean(lifeExp), max=max(lifeExp))
```


# finding the gdp

```{r}
total_gdp_by_country <- gapminder_data %>%
  mutate(gdp = pop * gdpPercap)

population_in_millions <- gapminder_data %>%
  mutate(popInMillions = pop / 1000000)

gapminder_data %>%
  select(- pop, - gdpPercap)
```


# ------------------------------------------

# filtering for only the Americas in the year 2007

```{r}
gapminder_data_2007_Americas <- read_csv("gapminder_data.csv") %>%
  filter(year == 2007 & continent == "Americas") %>%
  select(- year, - continent)
```


# ------------------------------------------

# Adding in CO2 Emissions

```{r}
CO2_emissions <- read_csv("co2-un-data.csv", skip = 2,
                          col_names= c("region", "country", "year", "series", "value", "footnotes", "source"))

co2_emissions_dirty <- CO2_emissions %>%
  select(country, year, series, value) 


co2_emissions_dirty <- CO2_emissions %>%
  select(country, year, series, value) %>%
  mutate(series = recode(series, "Emissions (thousand metric tons of carbon dioxide)" = "total_emissions",
                         "Emissions per capita (metric tons of carbon dioxide)" = "per_capita_emissions"))

co2_emissions_dirty %>%
  select(country, year, series, value) %>%
  mutate(series = recode(series, "Emissions (thousand metric tons of carbon dioxide)" = "total",
                         "Emissions per capita (metric tons of carbon dioxide)" = "per_capita")) %>%
  pivot_wider(names_from=series, values_from=value) %>%
  count(year)
```

# separating by year and emission per capital

```{r}
co2_emissions_dirty %>%
  select(country, year, series, value) %>%
  mutate(series = recode(series, "Emissions (thousand metric tons of carbon dioxide)" = "total",
                         "Emissions per capita (metric tons of carbon dioxide)" = "per_capita")) %>%
  pivot_wider(names_from=series, values_from=value) %>%
  count(year)
```


# filtered for the year 2005

```{r}
co2_emissions_dirty %>%
  select(country, year, series, value) %>%
  mutate(series = recode(series, "Emissions (thousand metric tons of carbon dioxide)" = "total",
                         "Emissions per capita (metric tons of carbon dioxide)" = "per_capita")) %>%
  pivot_wider(names_from=series, values_from=value) %>%
  filter(year == 2005) %>%
  count(year)
```


# assigning variable

```{r}
CO2_emmissions_2.0 <- 
  co2_emissions_dirty %>%
  select(country, year, series, value) %>%
  mutate(series = recode(series, "Emissions (thousand metric tons of carbon dioxide)" = "total",
                         "Emissions per capita (metric tons of carbon dioxide)" = "per_capita")) %>%
  pivot_wider(names_from=series, values_from=value) %>%
  count(year)
```

#------------------------------

# merging data frames

```{r}
CO2_emissions <- read_csv("co2-un-data.csv", skip=2,
                          col_names=c("region", "country", "year",
                                      "series", "value", "footnotes", "source")) %>%
  select(country, year, series, value) %>%
  mutate(series = recode(series, "Emissions (thousand metric tons of carbon dioxide)" = "total",
                         "Emissions per capita (metric tons of carbon dioxide)" = "per_capita")) %>%
  pivot_wider(names_from=series, values_from=value) %>%
  filter(year==2005) %>%
  select(-year) %>%
  mutate(country=recode(country,
                        "Bolivia (Plurin. State of)" = "Bolivia",
                        "United States of America" = "United States",
                        "Venezuela (Boliv. Rep. of)" = "Venezuela")
  )

anti_join(gapminder_data, CO2_emissions, by="country")

gapminder_data <- read_csv("gapminder_data.csv") %>%
  filter(year == 2007 & continent == "Americas") %>%
  select(-year, -continent) %>%
  mutate(country = recode(country, "Puerto Rico" = "United States")) %>%
  group_by(country) %>%
  summarize(lifeExp = sum(lifeExp * pop)/sum(pop),
            gdpPercap = sum(gdpPercap * pop)/sum(pop),
            pop = sum(pop)
  )

anti_join(gapminder_data, CO2_emissions, by="country")

gapminder_CO2 <- inner_join(gapminder_data, CO2_emissions, by="country")

gapminder_CO2 %>%  
  mutate(region = if_else(country == "Canada" | country == "United States" | country == "Mexico", "north", "south"))  

gapminder_and_CO2 <- gapminder_CO2 %>%  
  mutate(region = if_else(country == "Canada" | country == "United States" | country == "Mexico", "north", "south"))  
```


#-----------------------

# plotting my data

```{r}
ggplot(gapminder_and_CO2, aes(x=gdpPercap, y=per_capita)) +
  geom_point() +
  labs(x="GDP (per capita)",
       y="CO2 emitted (per capita)",
       title="There is a strong association between a nation's GDP \nand the amount of CO2 it produces"
  ) +
  geom_smooth(method =lm)


gapminder_and_CO2 %>%
  mutate(region = if_else(country == "Canada" | 
                            country == "United States" | 
                            country == "Mexico", 
                          "north", "south"))

gapminder_and_CO2 %>%
  mutate(region = if_else(country == "Canada" |
                            country == "United States" |
                            country == "Mexico", "north", "south")) %>%
  group_by(region) %>%
  summarize(sumtotal = sum(total),
            sumpop = sum(pop))
```


ggplot(gapminder_and_CO2, aes(x=gdpPercap, y=per_capita)) +
  geom_point() +
  labs(x="GDP (per capita)",
       y="CO2 emitted (per capita)",
       title="There is a strong association between a nation's GDP \nand the amount of CO2 it produces"
  ) +
  geom_smooth(method =lm)


gapminder_and_CO2 %>%
  mutate(region = if_else(country == "Canada" | 
                            country == "United States" | 
                            country == "Mexico", 
                          "north", "south"))

gapminder_and_CO2 %>%
  mutate(region = if_else(country == "Canada" |
                            country == "United States" |
                            country == "Mexico", "north", "south")) %>%
  group_by(region) %>%
  summarize(sumtotal = sum(total),
            sumpop = sum(pop))